# Parsing OPTA f24 XML files into R

I can't remember where I found this solution but this is not my original work, if you are reading this and you are the orginal author then please let me know. 

This script allows us to import a OPTA f24 file and twist it into a manageable dataframe object. 

First up is to load the libraries, if you have not installed them yet do this first. 

```{r}
library(XML)
library(plyr)
library(reshape)
```

Make sure the f24 file if in your current working folder and then set it as a variable 'fnm'. 

```{r}
fnm <- "yourf24filename.xml"
```

Next we setup a function that helps to collate all of the events at the end of the process 

```{r}
#utility function
grabAll <- function(XML.parsed, field){
  parse.field <- xpathSApply(XML.parsed, paste("//", field, "[@*]", sep=""))
  results <- t(sapply(parse.field, function(x) xmlAttrs(x)))
  if(typeof(results)=="list"){
    do.call(rbind.fill, lapply(lapply(results, t), data.frame, stringsAsFactors=F))
  } else {
    as.data.frame(results, stringsAsFactors=F)
  }
}
```

Next we take the XML work out its structure and chop it up, stick it together in different segments and tidy it up a little. 

```{r}
#XML Parsing
pbpParse <- xmlInternalTreeParse(fnm)
eventInfo <- grabAll(pbpParse, "Event")
eventParse <- xpathSApply(pbpParse, "//Event")
NInfo <- sapply(eventParse, function(x) sum(names(xmlChildren(x)) == "Q"))
QInfo <- grabAll(pbpParse, "Q")
EventsExpanded <- as.data.frame(lapply(eventInfo[,1:2], function(x) rep(x, NInfo)), stringsAsFactors=F)
QInfo <- cbind(EventsExpanded, QInfo)
names(QInfo)[c(1,3)] <- c("Eid", "Qid")
QInfo$value <- ifelse(is.na(QInfo$value),-1, QInfo$value)
Qual <- cast(QInfo, Eid ~ qualifier_id)
```

Finally, we merge all the different segments together to create a final 'events' dataframe. Just to check it worked we print the head of the data.frame.

```{r}
events <- merge(eventInfo, Qual, by.x="id", by.y="Eid", all.x=T)
head(events)
```

Some values should be numeric but currently are in character format, so we change this as the final part of our parsing. 

```{r}
events$min <- as.numeric(as.character(events$min))
events$sec <- as.numeric(as.character(events$sec))
events$x <- as.numeric(as.character(events$x))
events$y <- as.numeric(as.character(events$y))
```

Bingo, the events are ready to use! 

## Full Code 

```{r}
library(XML)
library(plyr)
library(reshape)

fnm <- "yourf24filename.xml"

#utility function
grabAll <- function(XML.parsed, field){
  parse.field <- xpathSApply(XML.parsed, paste("//", field, "[@*]", sep=""))
  results <- t(sapply(parse.field, function(x) xmlAttrs(x)))
  if(typeof(results)=="list"){
    do.call(rbind.fill, lapply(lapply(results, t), data.frame, stringsAsFactors=F))
  } else {
    as.data.frame(results, stringsAsFactors=F)
  }
}

#XML Parsing
pbpParse <- xmlInternalTreeParse(fnm)
eventInfo <- grabAll(pbpParse, "Event")
eventParse <- xpathSApply(pbpParse, "//Event")
NInfo <- sapply(eventParse, function(x) sum(names(xmlChildren(x)) == "Q"))
QInfo <- grabAll(pbpParse, "Q")
EventsExpanded <- as.data.frame(lapply(eventInfo[,1:2], function(x) rep(x, NInfo)), stringsAsFactors=F)
QInfo <- cbind(EventsExpanded, QInfo)
names(QInfo)[c(1,3)] <- c("Eid", "Qid")
QInfo$value <- ifelse(is.na(QInfo$value),-1, QInfo$value)
Qual <- cast(QInfo, Eid ~ qualifier_id)

#final data set
events <- merge(eventInfo, Qual, by.x="id", by.y="Eid", all.x=T)
events$min <- as.numeric(as.character(events$min))
events$sec <- as.numeric(as.character(events$sec))
events$x <- as.numeric(as.character(events$x))
events$y <- as.numeric(as.character(events$y))

```
